Epistemic Semantics in Guarded String Models

### 2020-06-17 tex/draft.tex
update atoms

Kleene Algebra with Tests is an algebraic theory that is defined by
equations and inequalities, which a has model classes including
guarded string models, relational models, finite models, and matrix
models.  This paper focuses on defining a family of concrete guarded
string algebras, the elements of which are sets of guarded strings.
Definitions and notation mostly follow \cite{kozen2001automata}.
Additional syntax and semantics is included to model multi-agent
epistemic semantics.  Guarded strings over a finite alphabet $E$ are
like ordinary strings, but with truth assignments to a set $T$ of
primitive propositions (primitive tests) alternating with the symbols
from $E$.  In the algebra described by Figure \ref{epikfigure}, $E$ is the set of
events $\{a_1,a_0,b_1,b_0\}$, and in the elevator example, $\{u,d\}$.
In the elevator example, $T$ is $\{p,q\}$, and in the coin example it
is $\{h,t\}$.  We write truth assigments to the primitive propositions
that observe the constraint as Boolean vectors. In the coin example,
we get the vectors 01, 10 and in the elevator example, the vectors 00,
01, 10, 11.  Guarded strings are strings of events, alternating with
such vectors, and starting and ending with vectors.  \ex{gs1} gives
the encoding as guarded strings of the worlds in \ex{elevator1} and
\ex{coin1}.

\begin{example}   \label{gs1}
\begin{tabular}[t]{ll}
$v_1$ & $00 u 01 u 10 d 01 u 10$ \\
$v_2$ & $00 u 01 u 10 u 11$ \\   
$w_1$ & $01$ \\
$w_2$ & $01 a_0 01$ \\  
$w_3$ & $01 a_0 01 b_0 01$ \\
\end{tabular}
\end{example}

@@

\begin{example} \label{fusion1}
\hspace{-2mm}\( 00 u 01 u 10 d 01 \cdot 01 u 10 = 00 u 01 u 10 d 01 u 10 \)

\hspace{-2mm}\( 00 u 01 u 10 d 01 \cdot 10 u 11 = \mbox{\em{undefined}} \)

\hspace{-2mm}\(01 \cdot 01 a_0 01 = 01 a_0 01 \)

\hspace{-2mm}\(01 \cdot 10 a_1 10 = \mbox{\em{undefined}} \)
\end{example}

Rather than guarded strings, elements of a guarded string model for
KAT are sets of guarded strings.  In the application, these elements
have the interpretation of propositions (sets of possible worlds)
and/or event types.  An event such as $u$ in the guarded string model
corresponds to the set of guarded strings where the bare event is
flanked by compatible truth assigments, as defined in an Epik program
by the event declaration and the constraint declaration.  These have
the status of event types, in that they can ``happen'' in different
possible worlds. Here happening corresponds to incrementing the world
with some compatible element of the event type.  Where $e$ is a bare
event symbol, $\llbracket e \rrbracket^{\varphi,P}$ is the corresponding
set of guarded strings, consisting of the bare event decorated with
compatible truth assignments denoted by $P$ that are admitted by the
world-constraints $\varphi$.  See this formally in
Figure~\ref{fig:interp}, and the examples in \ex{events1}.
% u (!q!p):(!qp) + (!qp):(q!p) + (q!p):(qp) 173
% d (!qp):(!q!p) + (q!p):(!qp) + (qp):(q!p)

\begin{example} \label{events1}
\(
\begin{array}[t]{l|c}
  e & \llbracket e \rrbracket^{\varphi, P} \\ \hline  
  u & \{ 00u01, 01u10, 10u11\}\\
  d & \{ 01d00, 10d01, 11d10\}\\
  a_1 & \{10 a_1 10 \}\\
  a_0 & \{01 a_0 01\}\\
  b_1 & \{10 b_1 10 \}\\
  b_0 & \{01 b_0 01 \}\\
\end{array}
\)
\end{example}

>>>>

Kleene Algebra with Tests is an algebraic theory that is defined by
equations and inequalities, which a has model classes including
guarded string models, relational models, finite models, and matrix
models.  This paper focuses on defining a family of concrete guarded
string algebras, the elements of which are sets of guarded strings.
Definitions and notation mostly follow \cite{kozen2001automata}.
Additional syntax and semantics is included to model multi-agent
epistemic semantics.  Guarded strings over a finite alphabet $E$ are
like ordinary strings, but with truth assignments to a set $T$ of
primitive propositions (primitive tests) alternating with the symbols
from $E$. Following \cite{kozen2001automata}, we call these truth
assignments \emph{atoms}; the set of atoms is notated $\mathcal A$.
In the algebra described by Figure \ref{epikfigure}, $E$ is the set of
events $\{a_1,a_0,b_1,b_0\}$, and in the elevator example, $\{u,d\}$.
In the elevator example, $T$ is $\{p,q\}$, and in the coin example it
is $\{h,t\}$.  We write atoms as the boolean formulae that uniquely
specify the truth assignment. In the elevator example, we have atoms
the vectors $\bar q \bar p$, $\bar q p$, $q \bar p$, $q p$.  In the
coin example, we get $h\bar t$ and $\bar h t$, as well as the bogus
atoms $\bar h \bar t$ and $ht$!  To preclude bogosity, the
\texttt{constraint} declaration in Figure~\ref{epikfigure} describes
the set of \emph{feasible atoms}, which we parse into a formula
$\varphi$ (if there are multiple \texttt{constraint declarations},
$\varphi$ is the conjunction of the parsed formulae). For a formula
$\phi$, the set of feasible atoms is $\mathcal A_\phi$. In the coin
example, the constraint $\varphi$ is $h\bar t + \bar h t$. Since $\bar
h t \models \varphi$ and $h \bar t \models \varphi$, they are
permitted feasible atoms, and since $\bar h \bar t$ and $h t$ do not
satisfy $\phi$ they are not permitted.

Guarded strings are strings of events, alternating with such atoms,
and starting and ending with atoms.  \ex{gs1} gives the encoding as
guarded strings of the worlds in \ex{elevator1} and \ex{coin1}.

\begin{example}   \label{gs1}
\begin{tabular}[t]{ll}
$v_1$ & $\bar p \bar q ~u~ \bar p q ~u~ q \bar p ~d~ \bar p q ~u~ q \bar p$ \\
$v_2$ & $\bar p \bar q ~u~ \bar p q ~u~ p\bar q ~u~ pq$ \\   
$w_1$ & $\bar h t$ \\
$w_2$ & $\bar h t ~a_0~ \bar h t$ \\  
$w_3$ & $\bar h t ~a_0~ \bar h t ~b_0~ \bar h t$ \\
\end{tabular}
\end{example}

@@

\begin{example} \label{fusion1}
\hspace{-2mm}\( \bar p q ~u~ \bar p q ~u~ p \bar q ~d~ \bar p q \cdot \bar p q ~u~ p \bar q = \bar p q ~u~ \bar p q ~u~ p \bar q ~d~ \bar p q ~u~ p \bar q \)

\hspace{-2mm}\( \bar p \bar q ~u~ \bar p q ~u~ p \bar q ~d~ \bar p q \cdot p \bar q ~u~ pq = \mbox{\em{undefined}} \)

\hspace{-2mm}\(\bar h t \cdot \bar h t ~a_0~ \bar h t = \bar h t ~a_0~ \bar h t \)

\hspace{-2mm}\(\bar h t \cdot h \bar t ~a_1~ h \bar t = \mbox{\em{undefined}} \)
\end{example}

Rather than individual guarded strings, elements of a guarded string
model for KAT are sets of guarded strings.  In the application, these
elements have the interpretation of propositions (sets of possible
worlds) and/or event types.  An event such as $u$ in the guarded
string model corresponds to the set of guarded strings where the bare
event is flanked by compatible atoms. In standard KATs, this is  as defined in an Epik
program by the event declaration and the constraint declaration.
These have the status of event types, in that they can ``happen'' in
different possible worlds. Here happening corresponds to incrementing
the world with some compatible element of the event type.  Where $e$
is a bare event symbol, $\llbracket e \rrbracket^{\varphi,P}$ is the
corresponding set of guarded strings, consisting of the bare event
decorated with compatible truth assignments denoted by $P$ that are
admitted by the world-constraints $\varphi$.  See this formally in
Figure~\ref{fig:interp}, and the examples in \ex{events1}.
% u (!q!p):(!qp) + (!qp):(q!p) + (q!p):(qp) 173
% d (!qp):(!q!p) + (q!p):(!qp) + (qp):(q!p)

\begin{example} \label{events1}
\(
\begin{array}[t]{l|c}
  e & \llbracket e \rrbracket^{\varphi, P} \\ \hline  
  u & \{ \bar p \bar q ~u~ \bar p q, \bar p q ~u ~p \bar q,  p \bar q ~u~ pq\}\\
  d & \{ \bar p q ~d~ \bar p \bar q, p \bar q~d~\bar p q, p q~d~ p \bar q\}\\
  a_1 & \{h \bar t ~a_1~ h \bar t \}\\
  a_0 & \{t \bar h ~a_0~ t \bar h\}\\
  b_1 & \{h\bar t ~b_1~ h \bar t \}\\
  b_0 & \{t \bar h ~b_0~ t \bar h \}\\
\end{array}
\)
\end{example}
